\textbf{\sc Deliberative Autonomy for Marine Robotics}

### 2012-01-30 trunk/europa.tex
checkpoint. more europa content.

TODO: motivate with a few examples of problems that can be solved in CSP, Scheduling and Planning.

We begin with a brief overview of Constraint Programming.

@@

Constraint Programming (CP) is a discipline that provides a generic framework for representing, solving and making logical inference on constraints. A complete treatment of this discipline is given in [MAR98] and [APT03] and very concise introductions are provided in [BAR99] and [LUS01].

	A constraint programming problem consists of a set of variables V={x1,..,xN}, where each variable takes values from a domain d1,..,dN. Domains are most often discrete and finite, but there are numerous CP implementations associated with continuous domains. Given a defined conjunctive set of constraints on the variables: C={c1(x1,..,xN), ..., cK(x1,..,xN)}, the objective is to find one or more value assignments to V where all constraints are satisfied.

@@

@@

@@

	\item Extension Modules bundled with EUROPA: Temporal Network, Resources, NDDL, ANML.

@@

TODO: introduce model for Shopping or Rover example, link to architectural components.
explain variable, object, predicate and action types and instances, constraints, goals and facts.

@@

TODO: explain how inference is used to detect perform bounds propagation, domain reduction and to detect constraint violations.
- overview of the types of flaws and violations EUROPA detects.

@@

>>>>

Let's look at some of the problems that can be tackled using EUROPA.

Constraint Satisfaction: 
In the N-Queens problem, chess queens must be placed on an NxN chessboard in a way that there are no queens attacking each other.

Below is an example of a random positioning of queens on a NxN chessboard. Queens in violation of the non-attack constraint are highlighted in red.
TODO: include figure.

If we define NxN variables Qrc, r E [1,N], c E[1,N] , 
Qrc = 1 if Cell r,c in he chessboard is occupied by a Queen, 0 otherwise.

then we have to satisfy the following constraints:
Sum(Qrc) = 1 ForAll r (Only one Queen per row)
Sum(Qrc) = 1 ForAll c (Only one Queen per column)

Sum(Qr+i,c+i) = 1
Sum(Qr-1,c+i) = 1 (only on Queen on each diagonal)

We can solve the problem by finding assignments for all variables Qrc that satisfy the constraints above.

Below is a solution found by EUROPA using that formulation and a specialized search procedure.

TODO: include figure

Scheduling:
In the Resource Constrained Project Scheduling Problem (RCPSP), a project consisting of a set of activities must be scheduled in a way that satisfies minimum and/or maximum temporal separation constraints among them, as well as respects fixed limits on the availability of resources required to perform each activity. In addition to satisfying temporal and resource constraints, it is common for the user to want to minimize makespan so that the entire project is finished as early as possible.

Below is an example of a a solution provided by EUROPA for an RCPSP instance with 10 activities, 5 resources and 30 temporal constraints.
TODO: include figure

Planning:
In the Shopping Agent Problem (TODO: ref to Norvig) an agent needs to purchase a set of products that are available at specific locations (supermarket, hardware store, etc), the agent  is subject to temporal (must complete tasks by specific deadlines) and resource (fuel, carrying capacity, etc) constraints.

Below is a solution produced by EUROPA for a simple problem instance where the shopping agent needs to buy Bananas, Milk and a Drill.

TODO: include figure

As we can see, a planning problem (what actions must be performed to achieve a goal) may embed a scheduling problem (in what order should those actions be performed), and both embed Constraint Satisfaction problems as temporal, resource and many other kinds of constraints must be respected in most real life problems.

These relationships between Planning, Scheduling and Constraint Satisfaction have been examined in detail in the literature (TODOL ref Frank et al on Planning vs. Scheduling), and they lead to EUROPA's use of constraint reasoning as its innermost building block. Next we will examine the main ideas from Constraint Satisfaction Programming, and from Constraint Based Planning and Scheduling that constitute EUROPA's theoretical underpinnings, then we dive into how EUROPA implements those ideas.

@@

Constraint Satisfaction Programming, also know as Constraint Programming (CP) is a discipline that provides a generic framework for representing, solving and making logical inference on constraints. A complete treatment of this discipline is given in [MAR98] and [APT03] and very concise introductions are provided in [BAR99] and [LUS01].

	A constraint programming problem consists of a set of variables V={x1,..,xN}, where each variable takes values from a domain d1,..,dN. Domains are most often discrete and finite, but there are numerous CP implementations associated with continuous domains. Given a defined conjunctive set of constraints on the variables: $C=\{c1(x1,..,xN), ..., cK(x1,..,xN)\}$, the objective is to find one or more value assignments to V where all constraints are satisfied.

@@

\subsection{Constraint-Based Attribute and Interval Planning}
\label{sec:europa:cp}

There has been 
State Variable representation

CSP encodings

Constraint Based and interval planning.

@@

	TODO: talk about the token network, move stuff from the Concepts section if necessary

@@

	\item Extension Modules bundled with EUROPA: Temporal Network, Resources, NDDL, ANML.
		\begin{enumerate}
			\item Temporal Network
			\item Resources
			\item NDDL
			\item ANML (TODO:maybe it's better to leave out?)
		\end{enumerate}

@@

Shopping model:

// Locations (Home, SuperMarket, etc.)
class Location {
  string name;
  Location(string _name){
    name = _name;
  }
}

// Products (Milk, Banana, etc.) and
class Product {
  string name;
  Product(string _name) {
    name = _name;
  }
}

// ProductLocations (Banana can be found at SuperMarket, for example)
class ProductLocation {
  Location location;
  Product product;

  ProductLocation(Location _location, Product _product){
    location = _location;
    product = _product;
  }
}

// Use built-in Timeline functionality to enforce that an agent:
// a) Can't be at more than one place at a time.
// b) Can't Go more than one place at a time.
// c) Can't Go somewhere and be At somewhere at the same time.
class AgentLocation extends Timeline{
  predicate At {
    Location loc;
  }

  predicate Going {
    Location from;
    Location to;
  }
}

// In addition to having a location timeline, the agent can buy and own things.  
// Note that the actions and the location predicates can be concurrent 
// so they can't be on the same Timeline
class Agent {
  AgentLocation location;

  Agent() {
    location = new AgentLocation();
  }

  action Buy {
    Product product;
  }

  action Go {
    Location from;
    Location to;
  }

  predicate Own {
  	Product product;
  }
}

// Define the rules for our actions:
Agent::Go {
  met_by(condition object.location.At origin);
  eq(from, origin.loc);

  equals(effect object.location.Going going);
  eq(going.from, from);
  eq(going.to, to);

  meets(effect object.location.At destination);
  eq(to, destination.loc);
}

Agent::Buy {
  // A Buy takes 10 time units
  eq(10, duration);

  // initialized to all locations
  ProductLocation possibleStores;

  // limit possibleStores variables to ones that provide what we need to buy
  eq(product, possibleStores.product);

  // We must be At a location during our Buy, and that location must have the
  // product we want available:
  contained_by(condition object.location.At currLocation);
  eq(currLocation.loc, possibleStores.location);

  starts(effect Own purchase);
  eq(purchase.product,product);
}

And a problem instance:

// Allocate instances
Location Home = new Location("Home");
Location SuperMarket = new Location("SuperMarket");
Location HardwareStore = new Location("HardwareStore");

Product Banana = new Product("Banana");
Product Milk = new Product("Milk");
Product Drill = new Product("Drill");

ProductLocation bananaLocation = new ProductLocation(SuperMarket, Banana);
ProductLocation milkLocation = new ProductLocation(SuperMarket, Milk);
ProductLocation drillLocation = new ProductLocation(HardwareStore, Drill);

Agent agent = new Agent();

// Indicate that the database is closed - no new objects can be created
close();

// We start the day at Home:
fact(agent.location.At atHomeForBreakfast);
atHomeForBreakfast.loc.specify(Home);

// Goals for all of the agent's needs, buy if needed
goal(agent.Own gotMilk);
gotMilk.product.specify(Milk);
goal(agent.Own gotBanana);
gotBanana.product.specify(Banana);
goal(agent.Own gotDrill);
gotDrill.product.specify(Drill);

// Make sure agent is home for dinner
goal(agent.location.At atHomeForDinner);
atHomeForDinner.loc.specify(Home);

// Agent has all day to satisfy goals:
gotMilk after atHomeForBreakfast;
gotMilk before atHomeForDinner;

gotBanana after atHomeForBreakfast;
gotBanana before atHomeForDinner;

gotDrill after atHomeForBreakfast;
gotDrill before atHomeForDinner;

// Force things to happen within our planning horizon:
int horizonStart = 0;
int horizonEnd = 100;
leq(horizonStart, atHomeForBreakfast.start);
leq(atHomeForDinner.end, horizonEnd);

TODO: introduce model for Shopping link to architectural components.
explain variable, object, predicate and action types and instances, constraints, goals and facts.

@@

TODO: explain how inference is used to detect perform bounds propagation, domain reduction and to detect constraint violations.
- talk about AC3 implementation and extension mechanisms
- talk about europa's constraint library
- talk about Temporal Network
- talk about Resource Envelopes.

@@

TODO: expand a little more on how heuristics can be implemented by applying ordering/filtering for chooseFlaw and makeDecisionPoint
examples:
- deal with earliest flaws first
- deal with planning (open condition flaws) before scheduling (threats) ones
